---
output: github_document
---

# chessR

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/JaseZiv/chessR.svg?branch=master)](https://travis-ci.org/JaseZiv/chessR)
<!-- badges: end -->

## Overview

This package is designed to allow users to extract their game data from the popular online chess platform [chess.com](https://www.chess.com/).

The website offers a very convenient set of APIs to be able to access and documentation to these can be found [here](https://www.chess.com/news/view/published-data-api).


## Installation

You can install the chessR package from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/chessR")
```


## Functions

There are a number of functions available in this package, both for extracting data and also for analysing and visualising data.

### Extraction

The following extraction functions will get json formatted game data and clean and convert it to a data frame. The games that will be returned are from the following time classes on chess.com:

* "blitz";
* "bullet";
* "daily"; and
* "rapid"

#### Raw Data

The `get_raw_game_data()` function can take either a single, or multiple usernames. It will output a data frame with all the games played by that user.

```{r get_raw, eval = FALSE}
# install.packages("devtools")
raw_chess <- get_raw_game_data("JaseZiv")
```


#### Analysis Data

The following function will extract the same data that the `get_raw_game_data()` function will, however this function will also include additional columns to make analysing data easier.

The function can be used either on a single player, or a character vector of multiple players.


```{r get_analysis, eval=FALSE}
chess_analysis_single <- get_game_data("JaseZiv")

chess_analysis_multiple <- get_game_data(c("JaseZiv", "elroch"))

```


#### Get Top Players' usernames

The below function allows the user to extract the top 50 leaders on the leaderboards for a number of different game types.

The game types include:

* "daily"
* "daily960""
* "live_rapid"
* "live_blitz"
* "live_bullet"
* "live_bughouse"
* "live_blitz960"
* "live_threecheck"
* "live_crazyhouse"
* "live_kingofthehill"
* "lessons"
* "tactics"

The usernames that are contained in the results can then be passed to `get_game_data` outlined above.

```{r get_leaders, eval=FALSE}
daily_leaders <- get_top50_leaderboard(game_type = "daily")
```
