---
output: github_document
---

# chessR

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/JaseZiv/chessR.svg?branch=master)](https://travis-ci.org/JaseZiv/chessR)
<!-- badges: end -->

## Overview

This package is designed to allow users to extract their game data from the popular online chess platform [chess.com](https://www.chess.com/).

The website offers a very convenient set of APIs to be able to access and documentation to these can be found [here](https://www.chess.com/news/view/published-data-api).


## Installation

You can install the chessR package from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/chessR")
```


## Usage

The functions available in this package are designed to enable the extraction of chess game data.

### Data Extraction

The functions detailed below relate to extracting data from chess.com


**Raw Game Data**

The `get_raw_game_data()` function can take either a single, or multiple usernames. It will output a data frame with all the games played by that user.

```{r get_raw, eval = FALSE}
# install.packages("devtools")
raw_chess <- get_raw_chessdotcom("JaseZiv")
```


**Analysis Data**

The following function will extract the same data that the `get_raw_chessdotcom()` function will, however this function will also include additional columns to make analysing data easier.

The function can be used either on a single player, or a character vector of multiple players.


```{r get_analysis, eval=FALSE}
chess_analysis_single <- get_game_data("JaseZiv")

chess_analysis_multiple <- get_game_data(c("JaseZiv", "elroch"))

```


**Get Top Players' usernames**

The below function allows the user to extract the top 50 leaders on the leaderboards for a number of different game types.

The game types include:

> *"daily", "daily960", "live_rapid", "live_blitz", "live_bullet", "live_bughouse", "live_blitz960", "live_threecheck", "live_crazyhouse", "live_kingofthehill", "lessons" and "tactics"*.

The usernames that are contained in the results can then be passed to `get_raw_chessdotcom()` outlined above.

```{r get_leaders, eval=FALSE}
daily_leaders <- get_top50_leaderboard(game_type = "daily")
```


For a detailed guide to using the package and the functions for analysis, see the package [vignette](https://jaseziv.github.io/chessR/articles/using_chessR_package.html)
